### 2023-01-15 paper/code/paper.tex
create proper gitignore + add paper

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Territory size, governance and economic performance: evidence from German dynasties}
\author{Elias Hadj Ammar}
\date{January 2023}

\begin{document}

\maketitle

Alternative title ideas: 
\begin{itemize}
    \item "Good riddance? What happens to states when dynasties go extinct" - catchy, but obscure, and also not what the paper is about.
    \item "What kind of ruler makes states prosper?" "The best kind of autocrat?"
    \item "How individual dynasties in the HRE promoted or hindered growth" - to the point, but detached from literature.
    \item Winners and losers
    \item Good riddance
    \item Thinly veiled dick joke
    \item Preview of results
\end{itemize}

\section{Introduction}

\end{document}