### 2023-01-15 build/code/exploration.Rmd
add R markdown for exploration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
```

## Ruler changes, family extinctions and territory outcomes

Questions I can think of:
- how are family extinctions different from regular successions?
- how can I get a list of these events from the data? How many are there?
- what kind of outcome data can I find on a city or territory level?

```{r data import}
territories <- read.csv("build/input/territories.csv")
summary(territories)
```

Is the variable `sovereign_number` the same as `ruler_id` in my data? The first thing to do is really really understand the data I have at my disposal. If the published data set doesn't have what I need, I would have to talk to Matthias and Prof. Cantoni about using the unpublished data.

First of all, drop a shitton of useless variables.

```{r variable selection}
territories <- territories %>%
  select(city_id, terr_id, sovereign_number, time_point, observation, point_id, 
         type_reign, type_change, city_status, foreign_rule)

View(territories)
```

From just viewing this, I can already see that this is not the data I want, at least not fully. 
These are city-year level observations of `terr_id`, that is, of ruling dynasty. So this - the already-published `territories` data - is basically a key to link `rulers` and `cities` data. I know that the `rulers` data exist, because I worked on them, and I am sure there is city-level data on outcomes like economic performance (wages, population, urbanisation) for the HRE out there.

So I need to 
- ask Prof. Cantoni for permission to use the `rulers` data
- identify and filter out family extinction events
- compile my own `outcomes` data set from other sources
- match it to the cities in the published `territories` data
- run a staggered diff-in-diff on that.