### 2023-01-16 build/code/exploration/data_structure.Rmd
add exploration and GetRulers.R

---
title: "Target for the data"
output: html_document
date: "2023-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(haven)
```

## What would I run the regressions on?

The regression would look something like this:

```{r cars}
reg1 <- lm(pop ~ city_fe + year_fe + treat*post + controls)
```

Ideally, I want a city-year level dataset with an entry for every year, based on the ruler dataset, not the territories dataset. For any given year, I want to know immediately not only the city's ruling lineage, but also the ruler. And then, for these city-year observations, I can add my outcome variables.

## Relational database approach

How do I do that? It seems to me that this kind of works like a relational database.
Each city-year observation must have non-missing lineage. Each lineage-year observation must have a non-missing ruler. That's how I could get there.

### Lineage for all city-year observations

But first let me check whether this is already in `cities_families.dta`.

```{r cities data}
cities <- read_dta('build/input/cities_families_1300_1918.dta')
summary(cities)
view(cities)
```

Okay, perfect. That file already contains the lineages for all cities for all years. At first glance it seems to be exactly what I would have had to do with the territories data - expand it so there's an observation for each year, based on the exact years in the commentary fields. That would have been a huge pain.

### Ruler for all lineage-year observations

The file `family_rulers_imputed.dta` is a list of ruler-level observations. These rulers happen to be in chronological order, but they're not lineage-year observations. Yet. How do I get them there?

```{text pseudo-code}
lineages <- expand(terr_id x year)
lineages <- mutate(ruler_id = case_when(
  year %in% start_reign:end_reign ~ ruler_id
))
```

Something like that may get the job done, but it seems computationally intensive and unelegant. Mathematically, what am I trying to do? I'm not trying to map lineage-year observations to rulers. I'm

Maybe do a function, or a dictionary. You input a terr_id and a year. It filters the rulers data down to the terr_id. Then it drops all rulers with reigns ending before or starting after that year. it returns the ruler of that lineage of that year.

Then `map` year-lineage entries to rulers with that function.