### 2023-01-16 build/code/exploration/data_structure.Rmd

---
title: "Target for the data"
output: html_document
date: "2023-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(haven)
```

## What would I run the regressions on?

The regression would look something like this:

```{r cars}
reg1 <- lm(pop ~ city_fe + year_fe + treat*post + treat*post*channel + controls)
```

In the end, `treat`x`post` will be 1 for all city-years after that city's extinction year. That's my intuition. But I haven't read up about staggered diff-in-diff yet.

Ideally, I want a city-year level dataset with an entry for every year, based on the ruler dataset, not the territories dataset. For any given year, I want to know immediately not only the city's ruling lineage, but also the ruler. And then, for these city-year observations, I can add my outcome variables.

## Relational database approach

How do I do that? It seems to me that this kind of works like a relational database.
Each city-year observation must have non-missing lineage. Each lineage-year observation must have a non-missing ruler. That's how I could get there.

### Lineage for all city-year observations

But first let me check whether this is already in `cities_families.dta`.

```{r cities data}
cities <- read_dta('build/input/cities_families_1300_1918.dta')
summary(cities)
view(cities)
```

Okay, perfect. That file already contains the lineages for all cities for all years. At first glance it seems to be exactly what I would have had to do with the territories data - expand it so there's an observation for each year, based on the exact years in the commentary fields. That would have been a huge pain.

### Ruler for all lineage-year observations

The file `family_rulers_imputed.dta` is a list of ruler-level observations. These rulers happen to be in chronological order, but they're not lineage-year observations. Yet. How do I get them there?

```{text pseudo-code}
lineages <- expand(terr_id x year)
lineages <- mutate(ruler_id = case_when(
  year %in% start_reign:end_reign ~ ruler_id
))
```

I did something like that in `RulersToTerritoryYears.R`. Surprisingly, only about one-third of territory-year observation have a corresponding ruler who was alive at the time. But I don't think that should be an issue. I should look at some descriptive stats.

```{r descriptive stats of rulers_linked}
rulers <- read.csv("build/temp/rulers_linked.csv")
print(rulers %>% group_by(terr_id) %>% summarise(years = n()) %>% summary())

```
The median dynasty in my data lasted 106 years. That seems plausible.

```{r more descriptive stats}
print(rulers %>% group_by(year) %>% summarise(territories = n()) %>% summary())
```
If I did it right, this means that there were on average 75 dynasties with rulers coexisting in the same year. Also plausible.

It would be interesting why the other 100,000 dynasty-year observations fell through.
Is it that some years (e.g. early years) are often missing rulers?
Is it that some dynasties (e.g. small ones) are often missing rulers?
Could there be an explanation other than data quality?
This is a potential source of bias; I should address it. Maybe limit the analysis to cities in territories larger than a certain threshold, and to recent centuries in which there are few missing links.

One more note: `rulers_linked.csv` essentially contains the personal history of all these dynasties. It contains the extinction events and, crucially, the last ruler of that dynasty, whose characteristics I can link to the event from elsewhere.
#