### 2023-01-17 build/code/exploration/PopData.Rmd
add population data (not cited yet)

---
title: "Europop Data (De Vries 1984)"
output: html_document
date: "2023-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(haven)
```

## First look at the population data

```{r first look}
devries <- read.csv("build/input/devries1984.csv")
bairoch <- read.csv("build/input/bairoch1988.csv")
bairoch %>% drop_na(population) %>% view()
```

Um. Okay, that's really not super granular. 50 is the coarsest I would have gone, and it is apparently also the finest I can go. But it will do. I just need to limit my analysis to only those extinction events with at least 50 or 100 years of stability afterwards.

## Extracting the cities I want (all in `cities`)

## Adding a population variable to the city-year observations