### 2023-02-23 analysis/code/SummaryStats.Rmd
build lineage size and extinctions

---
title: "SummaryStats"
output: html_document
date: "2023-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
```

## Import datasets

```{r import data}
ImportRulers <- function(){
  rulers <- read.csv("build/temp/rulers.csv")
  rulers <- rulers %>% 
    select(id, family, name, terr_id, birth_year, 
           death_year, start_reign, end_reign)
  return(rulers)
}

rulers <- ImportRulers()
```

## Summary statistics: lineages

```{r count rulers for each dynasty}
CountRulers <- function(rulers){
  last_rulers <- rulers %>% 
    group_by(terr_id) %>% 
    count()
}

ruler_counts <- CountRulers(rulers)
ruler_counts %>% View()
```