### 2023-03-01 analysis/code/DuplicateKeys.Rmd
#1 explore dupe keys issue

## Ensuring that city-year keys are unique

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(cowplot)
library(haven)
```

What I've established in Exploration.Rmd:

- In the current version of build.csv, there are 3708 rows that have duplicate city-year keys.
- It turns out that all these rows have `terr_id` "A1341" (Anhalt-Dessau, younger line).
- It turns out that exactly half of all observations with "A1341" are dupes.
- It turns out that the only things that vary for dupes are `X` (a key that was automatically generated in some join operation) and `count_diff` (the difference between `count_cities` and `lag(count_cities)`, which is calculated rowwise).
- It turns out that for each city-year, only the first dupe has the right `count_diff`.

It does seem like I could simply run distinct(city_id, year) on my current data build and get unique city-year keys without losing any information. But this could happen again, on a larger scale, for more territories, if I change the build code somehow. So I want to find out where the duplication happens - in the original cities data, or at any point in the build pipeline - and fix the bug properly.

```{r data import, message=FALSE}
cities_orig <- read_dta("build/input/cities_families_1300_1918.dta") %>% 
  select(city_id, year, terr_id)

cities <- read_csv("build/output/cities.csv", show_col_types = FALSE)
lineages <- read_csv("build/output/lineages.csv", show_col_types = FALSE)
build <- read_csv("build/output/build.csv", col_types = cols( # necessary because first 1000 are NA
  city="c", 
  real_wage="d", 
  welfare_ratio="d"
  ))
```

### Are there dupes in the original cities data?

```{r check cities_orig}
summary(cities_orig) # no missing terr_id for any city-year - good

cities_orig %>% filter(terr_id == "A1341") %>% count() # the dupes are not in the original file!

# double-check: yes, no duplicate keys in cities_orig.
cities_orig %>% count() - cities_orig %>% distinct(city_id, year) %>% count()
```

### Are there dupes in the output of CombineCityData.R?

No. As of now, the output of that file (cities.csv) has the same city-years as the original data. Also, it uses a left join to add wages to cities - I think this is inherently less problematic.

Just for good measure (and I can just re-run this when cities.csv changes):

```{r check cities.csv}
summary(cities) # yep, same as above

cities %>% filter(terr_id == "A1341") %>% count() # same as above

# no dupes in cities
cities %>% count() - cities %>% distinct(city_id, year) %>% count()
```

This means that the bug must be in CombineAll.R.

### Where in CombineAll.R does the duplication happen?

It must be in CombineTables because as I've seen, the AddSizeDiff already operates on data that contains duplicates. It's the inner join between the lineages and the cities data in CombineTables.

What if I just do a left join instead?

Expected result:

- a data frame that is the same length as cities (1.4 million obs. rather than 1.1 million)
- cities with `terr_id` that doesn't show up in the lineages data have NA values
- no duplicate city-year keys
- summary statistics identical to the previous build (after dropping `count_cities` = NA)

```{r try left join in CombineTables}
CombineTables <- function(cities, lineages){
  joined <- left_join(cities, lineages, by=c("terr_id", "year")) 
  return(joined)
}

new_build <- CombineTables(cities, lineages)
summary(new_build)
```

I did not get the expected result. Why?
I think it must be because "A1341" shows up twice in the lineage data.

### Are there dupes in the lineage data?

```{r check lineage data}
# Yes, there are dupes in lineages.csv:
lineages %>% count() - lineages %>% distinct(terr_id, year) %>% count()

# The 315 dupes in lineages.csv are 315 dupes of terr-years of A1341:
lineages %>% filter(terr_id == "A1341") %>% count()
lineages %>% filter(terr_id == "A1341") %>% distinct(terr_id, year) %>% count()

# The duplicate terr-year keys are full duplicates of rows:
lineages %>% select(-...1) %>% filter(terr_id == "A1341") %>% distinct() %>% count()
```

### Where do the dupes in the lineage data come from?

It must be CombineLineageData.R, and within that file it must be AddExtinctionDummies or AddLineageNames.

Old Hypothesis:
There is no reason to use an inner join in any of the functions there, and using a left join instead will solve the duplicates problem. EXCEPT if there are two last rulers (or two lineage names) for A1341.

New Hypothesis:
In the year Anhalt-Dessau went extinct, there were two rulers.

```{r check last_rulers.csv}
last_rulers <- read_csv("build/temp/last_rulers.csv")

last_rulers %>% filter(terr_id == "A1341") # HAH, I knew it.

# Very interesting - there are *two* lineages with two last rulers.
last_rulers %>% count() - last_rulers %>% distinct(terr_id) %>% count()

# Which ones:
anti_join(last_rulers, last_rulers %>% distinct(terr_id, .keep_all=TRUE), by="...1")

# The other one is of Wessex, terr_id "0W0079". Do we have cities in Wessex? No...
cities_orig %>% filter(terr_id == "0W0079")

# ... which is why this other duplicate extinction doesn't lead to dupes in the build.
```

Solution: modify LineageExtinctions.R so that last_rulers.csv has `terr_id` as a *unique* key. See if that fixes the problem.