### 2023-04-10 analysis/code/exploration/ExploreOutcomes.Rmd
#4 rework outcome plotting

## Exploration: Outcomes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/GitHub/BA')
library(tidyverse)
library(cowplot)
library(viridisLite)
```

### Data import and summary

```{r data import}
data <- read.csv("analysis/input/build.csv")
```

### List cities to plot

```{r}
ListCitiesNotNA <- function(data, outcome, window=1300:1918){
  ids <- data %>% 
    filter(year %in% window) %>% 
    drop_na({{outcome}}) %>% 
    distinct(city_id) %>% 
    .$city_id
  return(ids)
}

ListTopCities <- function(data, which=1:8, by="obs", window=1300:1918){
  cities_life <- data %>% 
    filter(year %in% window) %>% 
    group_by(city_id) %>% 
    summarise(
      lifetime_construction = sum(construction, na.rm=TRUE),
      observations = sum(!is.na(construction))
      )

  if (by == "obs"){
    top_ids <- cities_life %>% 
      arrange(desc(observations), desc(lifetime_construction))
  } else {
    top_ids <- cities_life %>% 
      arrange(desc(lifetime_construction), desc(observations))
  }

  return(top_ids$city_id[which])
}
```

```{r}
# city list for comparison
const_cities <- ListCitiesNotNA(data, construction)
allen_cities <- ListCitiesNotNA(data, real_wage)
both <- intersect(const_cities, allen_cities)

# city lists for exploring construction
set.seed(151101)
top9 <- ListTopCities(data, which=1:9)
top16 <- ListTopCities(data, which=1:16)
random9 <- sample(const_cities, 9)
random16 <- sample(const_cities, 16)

```

### Plotting functions

```{r}
PlotOutcomeLines <- function(data, citylist=NULL, outcome, colour_vec, window=1300:1918){
  if (length(citylist) == 0){
    city_list <- ListCitiesNotNA(data, {{outcome}}, window)
  }
  plot <- data %>% 
    filter(city_id %in% citylist & year %in% window) %>% 
    mutate(city_id = factor(city_id)) %>% 
    drop_na({{outcome}}) %>% 
    ggplot(aes(x=year, y={{outcome}}))+
    geom_line(aes(colour=city_id))+
    scale_color_manual("City", values=colour_vec)+
    scale_x_continuous(limits=c(min(window), max(window)))

  return(plot)
}

PlotHistograms <- function(data, citylist, binw=10){
  data |>
    filter(city_id %in% citylist) |> 
    mutate(city_id = as_factor(city_id)) |> 
    filter(construction > 0) |> 
    ggplot(aes(x=year))+
    geom_histogram(aes(y=after_stat(density)), binwidth = binw)+
    geom_density(aes(y=after_stat(density), colour="red"), show.legend = FALSE)+
    facet_wrap(~city_id)
}

PlotDensity <- function(data, citylist, colour_vec, window=1300:1918){
  data |>
    filter(city_id %in% citylist & year %in% window) |> 
    mutate(city_id = as_factor(city_id)) |> 
    filter(construction > 0) |> 
    ggplot(aes(x=year))+
    geom_density(aes(y=after_stat(density), colour=city_id))+
    scale_color_manual("City", values=colour_vec)+
    scale_x_continuous(limits=c(min(window), max(window)))
}

```

### Comparison to real wages

```{r plots for Allen cities}
colour_vec <- c("gold", "cornflowerblue", "coral")
names(colour_vec) <- allen_cities

wage_plot <- PlotOutcomeLines(data, citylist=both, real_wage, colour_vec, window=1400:1800)
wratio_plot <- PlotOutcomeLines(data, citylist=both, welfare_ratio, colour_vec, window=1400:1800)
const_plot <- PlotDensity(data, citylist = both, colour_vec, window=1400:1800)

plot_grid(wage_plot, wratio_plot, const_plot, align="v", axis="b", ncols=1)
ggsave("const_wages.png", device = "png", path="analysis/output/exploration")
```

What's striking and concerning is that they really don't match at all. Maybe I need to control for something? How would that work, though?

Now I want to see construction for some more cities, but not for multiple thousand.

```{r}

PlotHistograms(data, random16) +
  ggtitle("Construction activity: random cities")
PlotHistograms(data, top16) +
  ggtitle("Construction activity: cities with the most observations")

```

I want to see how the number of recorded events is related to
- a city's population in 1875
- a city's age