### 2023-06-21 paper/output/regressions/window_10y.tex

\begin{table}[htbp]
   \caption{\label{tab:window_10y} Differences in trends after switching, by switch type}
   \centering
   \begin{tabular}{lcccc}
      \tabularnewline \midrule \midrule
      Dependent Variables:                       & All construction & State   & Private               & Public goods\\  
      Model:                                     & (1)              & (2)     & (3)                   & (4)\\  
      \midrule
      \emph{Variables}\\
      Switch to another state                    & 0.01$^{*}$       & 0.003   & -0.004                & -0.008$^{**}$\\   
                                                 & (0.009)          & (0.004) & (0.003)               & (0.003)\\   
      Switch to another state $\times$ Conquest  & -0.008           & -0.01   & 0.003                 & 0.004\\   
                                                 & (0.01)           & (0.007) & (0.005)               & (0.005)\\   
      Switch to another state $\times$ Other     & -0.01            & -0.005  & 0.002                 & 0.01$^{**}$\\   
                                                 & (0.01)           & (0.005) & (0.004)               & (0.005)\\   
      Switching indicator                        & -0.03$^{***}$    & -0.004  & -0.005                & -0.010$^{**}$\\   
                                                 & (0.01)           & (0.005) & (0.004)               & (0.004)\\   
      \midrule
      \emph{Fixed-effects}\\
      City                                       & Yes              & Yes     & Yes                   & Yes\\  
      Period                                     & Yes              & Yes     & Yes                   & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations                               & 35,168           & 35,168  & 35,168                & 35,168\\  
      R$^2$                                      & 0.1770           & 0.0835  & 0.1278                & 0.0890\\  
      Within R$^2$                               & 0.0004           & 0.0001  & $8.13\times 10^{-5}$  & 0.0004\\  
      \midrule \midrule

   \end{tabular}

   \par \raggedright 
   Note: Table presents results of estimation equation \eqref{eq:baseline}. The switch type "Succession" is omitted as the  reference category. Yearly data was aggregated into periods of 10 years. Observations are at the city-period  level. Dependent variables are indicators that take the value 1 if  construction activity of the respective type was recorded. Standard errors are  clustered at the city level. *, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent  level, respectively.
\end{table}