### 2023-06-22 paper/output/regressions/SW22_replication_100y.tex
#28 proper SW22 results table

\begin{table}[htbp]
   \caption{\label{tab:SW22_replication_100y} Dynamic effects of switching}
   \centering
   \begin{tabular}{lcccc}
      \tabularnewline \midrule \midrule
      Dependent Variables:           & All construction & State      & Private       & Public goods\\  
      Model:                         & (1)              & (2)        & (3)           & (4)\\  
      \midrule
      \emph{Variables}\\
      Treat $\times$ Period $=$ 0    & 0.02             & -0.03      & 0.0006        & -0.04$^{**}$\\   
                                     & (0.02)           & (0.03)     & (0.02)        & (0.02)\\   
      Treat $\times$ Period $=$ 100  & 0.09$^{***}$     & 0.02       & 0.01          & -0.008\\   
                                     & (0.03)           & (0.03)     & (0.03)        & (0.03)\\   
      Treat $\times$ Period $=$ 200  & 0.08$^{***}$     & -0.02      & -0.10$^{***}$ & -0.06$^{*}$\\   
                                     & (0.03)           & (0.04)     & (0.03)        & (0.04)\\   
      Switching indicator            & 0.02             & 0.06$^{*}$ & 0.009         & -0.02\\   
                                     & (0.03)           & (0.03)     & (0.02)        & (0.03)\\   
      \midrule
      \emph{Fixed-effects}\\
      City                           & Yes              & Yes        & Yes           & Yes\\  
      Period                         & Yes              & Yes        & Yes           & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations                   & 2,955            & 2,955      & 2,955         & 2,955\\  
      R$^2$                          & 0.4303           & 0.4530     & 0.4832        & 0.4499\\  
      Within R$^2$                   & 0.0048           & 0.0047     & 0.0064        & 0.0027\\  
      \midrule \midrule

   \end{tabular}

   \par \raggedright 
   Note: Table presents results of estimation equation \eqref{eq:sw22}. Yearly data was aggregated into periods of 100 years. Observations are at the city-period  level. Dependent variables are indicators that take the value 1 if  construction activity of the respective type was recorded. Standard errors are  clustered at the city level. *, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent  level, respectively.
\end{table}