### 2023-06-22 paper/output/regressions/window_50y.tex

\begin{table}[htbp]
   \caption{\label{tab:window_50y} Differences in trends after switching, by switch type}
   \centering
   \begin{tabular}{lcccc}
      \tabularnewline \midrule \midrule
      Dependent Variables:                       & All construction & State         & Private & Public goods\\  
      Model:                                     & (1)              & (2)           & (3)     & (4)\\  
      \midrule
      \emph{Variables}\\
      Switch to another state                    & 0.02             & 0.02          & 0.00    & -0.04$^{***}$\\   
                                                 & (0.02)           & (0.02)        & (0.01)  & (0.01)\\   
      Switch to another state $\times$ Conquest  & 0.04             & -0.07$^{***}$ & 0.00    & 0.02\\   
                                                 & (0.04)           & (0.03)        & (0.02)  & (0.02)\\   
      Switch to another state $\times$ Other     & 0.01             & -0.02         & 0.01    & 0.04$^{**}$\\   
                                                 & (0.03)           & (0.02)        & (0.01)  & (0.02)\\   
      Switching indicator                        & -0.01            & 0.01          & -0.01   & 0.00\\   
                                                 & (0.02)           & (0.02)        & (0.01)  & (0.01)\\   
      \midrule
      \emph{Fixed-effects}\\
      City                                       & Yes              & Yes           & Yes     & Yes\\  
      Period                                     & Yes              & Yes           & Yes     & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations                               & 6,635            & 6,635         & 6,635   & 6,635\\  
      R$^2$                                      & 0.3553           & 0.2798        & 0.3257  & 0.2917\\  
      Within R$^2$                               & 0.0009           & 0.0012        & 0.0002  & 0.0011\\  
      \midrule \midrule

   \end{tabular}

   \par \raggedright 
   Note: Table presents results of estimation equation \eqref{eq:baseline}. The switch type "Succession" is omitted as the  reference category. Yearly data was aggregated into periods of 50 years. Observations are at the city-period  level. Dependent variables are indicators that take the value 1 if  construction activity of the respective type was recorded. Standard errors are  clustered at the city level. *, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent  level, respectively.
\end{table}