### 2023-06-22 paper/output/regressions/controls_50y.tex
#29 output controls table

\begin{table}[htbp]
   \caption{\label{tab:controls_50y} Heterogeneous effects - different controls}
   \centering
   \begin{tabular}{lcccc}
      \tabularnewline \midrule \midrule
      Dependent Variable: & \multicolumn{4}{c}{All construction}\\
      Model:                                     & (1)    & (2)    & (3)           & (4)\\  
      \midrule
      \emph{Variables}\\
      Switch to another state                    & 0.02   & 0.02   & -0.01         & 0.03\\   
                                                 & (0.03) & (0.02) & (0.05)        & (0.05)\\   
      Switch to another state $\times$ Conquest  & 0.04   & 0.04   & -0.15$^{***}$ & -0.14$^{**}$\\   
                                                 & (0.04) & (0.04) & (0.05)        & (0.06)\\   
      Switch to another state $\times$ Other     & 0.01   & 0.01   & 0.05          & 0.05\\   
                                                 & (0.03) & (0.03) & (0.05)        & (0.05)\\   
      Switching indicator                        & -0.01  &        & -0.09         &   \\   
                                                 & (0.02) &        & (0.06)        &   \\   
      Conflict                                   &        &        & 0.01          & 0.01\\   
                                                 &        &        & (0.03)        & (0.03)\\   
      \midrule
      \emph{Fixed-effects}\\
      City                                       & Yes    & Yes    & Yes           & Yes\\  
      Period                                     & Yes    & Yes    & Yes           & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations                               & 6,635  & 6,635  & 1,955         & 1,955\\  
      R$^2$                                      & 0.3553 & 0.3553 & 0.4245        & 0.4230\\  
      Within R$^2$                               & 0.0009 & 0.0008 & 0.0071        & 0.0045\\  
      \midrule \midrule

   \end{tabular}

   \par \raggedright 
   Note: Table presents results of estimation equation \eqref{eq:baseline} but using different control variables. The switch type  "Succession" is omitted as the reference category. Yearly data was aggregated into periods of 50 years. Observations are at the city-period level. Dependent variables are indicators  that take the value 1 if construction activity of the respective type was  recorded. Standard errors are clustered at the city level. *, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent  level, respectively.
\end{table}