### 2023-06-26 paper/output/regressions/controls_50y.tex
#29 include check in controls table

\begin{table}[htbp]
   \caption{\label{tab:controls_50y} Heterogeneous effects of switching: Robustness}
   \centering
   \begin{tabular}{lccccc}
      \tabularnewline \midrule \midrule
      Dependent Variable: & \multicolumn{5}{c}{All construction}\\
      Model:                                     & (1)    & (2)    & (3)           & (4)          & (5)\\  
      \midrule
      \emph{Variables}\\
      Switch to another state                    & 0.02   & 0.02   & -0.01         & 0.03         & -0.01\\   
                                                 & (0.03) & (0.02) & (0.05)        & (0.05)       & (0.05)\\   
      Switch to another state $\times$ Conquest  & 0.04   & 0.04   & -0.15$^{***}$ & -0.14$^{**}$ & -0.15$^{***}$\\   
                                                 & (0.04) & (0.04) & (0.05)        & (0.06)       & (0.05)\\   
      Switch to another state $\times$ Other     & 0.01   & 0.01   & 0.05          & 0.05         & 0.05\\   
                                                 & (0.03) & (0.03) & (0.05)        & (0.05)       & (0.05)\\   
      Switching indicator                        & -0.01  &        & -0.09         &              & -0.09\\   
                                                 & (0.02) &        & (0.06)        &              & (0.06)\\   
      Conflict                                   &        &        & 0.01          & 0.01         &   \\   
                                                 &        &        & (0.03)        & (0.03)       &   \\   
      \midrule
      \emph{Fixed-effects}\\
      City                                       & Yes    & Yes    & Yes           & Yes          & Yes\\  
      Period                                     & Yes    & Yes    & Yes           & Yes          & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations                               & 6,635  & 6,635  & 1,955         & 1,955        & 1,955\\  
      R$^2$                                      & 0.3553 & 0.3553 & 0.4245        & 0.4230       & 0.4244\\  
      Within R$^2$                               & 0.0009 & 0.0008 & 0.0071        & 0.0045       & 0.0069\\  
      \midrule \midrule

   \end{tabular}

   \par \raggedright 
   Note: Table presents results of estimation equation \eqref{eq:baseline} but using different control variables. Columns (3)-(5) drop observations with missing data on conflict incidents from the sample. Yearly data was aggregated into periods of 50 years. Observations are at the city-period level. The dependent variable is an  indicator that takes the value 1 if any construction activity was recorded.  Standard errors are clustered at the city level. *, **, and *** denote significance on the 10 percent, 5 percent, and 1 percent  level, respectively.
\end{table}